---
title: "visualization(Lecture 8).Rmd"
author: "Meher Walia"
date: "2026-10-06"
output: html_document
---

```{r setup, message = FALSE}
library(tidyverse)
library(ggridges)
library(p8105.datasets)

data("weather_df")
```
## Basic scatterplot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax))
```

# you do this if you dont want any points, it creates a plain scatterplot^^^

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

# do this when you want the points to be visible

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point()
```

```{r}
ggp_weather = 
  weather_df |>
  ggplot(aes(x = tmin, y = tmax)) 

ggp_weather + geom_point()
```

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name))
```
# do this if you want to add color to your scatterplot ^^

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5) +
  geom_smooth(se = FALSE)
```

#if you want a straight line down the middle to appear, use this^^

```{r}
weather_df |>
  ggplot(aes(x = tmin, y = tmax, color = name)) + 
  geom_point(alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

# use this if you want to split the graph into difdferent areas/panels^^

```{r}
weather_df |>
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) + 
  facet_grid(. ~ name)
```

# aes, reads the map, use prcp if you want precipitation aka big bubbles, tmax for maximum temperature and tmin for minimum^^

# Step 6: learning assessment
weather_df |>
  filter(name == "CentralPark_NY") |>
  mutate(
    tmax_fahr = tmax * (9 / 5) + 32,
    tmin_fahr = tmin * (9 / 5) + 32) |>
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point(alpha = .5) +
  geom_smooth(method = "lm", se = FALSE)
  
# weather_df |> — start with the full dataset (2,190 rows, all three stations) and pass it along.
filter(name == "CentralPark_NY") |> — keep only rows where the station is Central Park. Drops it to 730 rows. Note the double ==: one = means "assign," two means "is equal to."
mutate(...) |> — add two new columns. tmax_fahr = tmax * (9 / 5) + 32 takes each Celsius value and converts it. Same for tmin. The original Celsius columns stay, you've just added to them.
ggplot(aes(x = tmin_fahr, y = tmax_fahr)) + — start the plot, putting the new Fahrenheit columns on the axes. Note this switches from |> to +, because plot layers get added, not piped.
geom_point(alpha = .5) + — draw the dots. alpha = .5 makes them half-transparent so you can see where they pile up.
geom_smooth(method = "lm", se = FALSE) — draw a trend line. method = "lm" means linear model, so a straight line instead of a wiggly curve. se = FALSE hides the grey uncertainty band.

```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) + 
  geom_smooth(se = FALSE)
```

```{r}
weather_df |> 
  ggplot(aes(x = tmax, y = tmin)) + 
  geom_hex()
```

weather_df |> ggplot() + geom_point(aes(x = tmax, y = tmin), color = "blue")
weather_df |> ggplot() + geom_point(aes(x = tmax, y = tmin, color = "blue"))

# Inside aes() = map to a column. Outside aes() = set a fixed value.^^
geom_point(aes(color = name)) → color varies by station
geom_point(color = "blue") → every point is blue
If you get a legend you didn't want, you put a literal value inside aes() by mistake.

## Univariate plots

```{r}
weather_df |> 
  ggplot(aes(x = tmax)) + 
  geom_histogram()
```

```{r}
weather_df |> 
  ggplot(aes(x = tmax, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 2)
```

```{r}
weather_df |> 
  ggplot(aes(x = tmax, fill = name)) + 
  geom_density(alpha = .4, adjust = .5, color = "blue")
```

# Step 10: boxplots, violins, and ridges.
```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmax)) + 
  geom_boxplot()
```

```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmax)) + 
  geom_violin(aes(fill = name), alpha = .5) + 
  stat_summary(fun = "median", color = "blue")
```

```{r}
weather_df |> 
  ggplot(aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85)
```

#Learning assessment

# histogram
weather_df |> 
  ggplot(aes(x = prcp, fill = name)) + 
  geom_histogram(position = "dodge", binwidth = 10)

# density
weather_df |> 
  ggplot(aes(x = prcp)) + 
  geom_density(aes(fill = name), alpha = .5)

# boxplot
weather_df |> 
  ggplot(aes(y = prcp, x = name)) + 
  geom_boxplot()

# violin
weather_df |> 
  ggplot(aes(x = name, y = prcp)) + 
  geom_violin(aes(fill = name), alpha = .5)

# ridge, dropping zero-rain days so the shape is visible
weather_df |> 
  filter(prcp > 0) |> 
  ggplot(aes(x = prcp, y = name)) + 
  geom_density_ridges(scale = .85)
  
## Saving and embedding plots

```{r}
ggp_weather = 
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax)) + 
  geom_point(aes(color = name), alpha = .5)

ggsave("ggp_weather.pdf", ggp_weather, width = 8, height = 5)
```

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

```{r setup2, message = FALSE}
library(tidyverse)
library(ggridges)
library(p8105.datasets)

data("weather_df")

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```